Grocery stores in Canada

### 2024-11-14 paper/paper.qmd
fixed discussions

---
title: "Grocery stores in Canada"
subtitle: "What vendors hold the most items?"
author: 
  - Hyunje Park, Charlie Zhang
thanks: "Code and data are available at: https://github.com/davidpxrk/cost-of-living-toronto"
date: today
date-format: long
abstract: "This paper analyzes the Canadian Grocery Price data from Canadian grocery vendors, which was gathered by Jacob Filipp and hosted publicly by Project Hammer. It uses information on items, and the name of vendors to determine how many vendors hold a unique grocery item. It was discovered that vendors such as Loblaws, Metro and Walmart held the most unique items at around 20,000 items, while international-focused vendors such as Galleria and T&T held the least amount of items below 10,000.
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The `product` table contained 9 variables; most notably `id` and `vendor`. These two variables were important in this analysis, which captured the information on the product id, and the vendor that item was sold at. Data manipulation was conducted through `SQL` [@citeSQL], to select the variables `id`, `vendor`, and find the number of unique items each vendor sold. This information was exported as a `.csv` file, and the data analysis was conducted through `R Programming Language` [@citeR], `knitr` [@citeknitr],.

@@

The graph above show that vendors such as Metro and Walmart far exceeded the number of products compared to vendors such as Galleria, or T&T, highlighting a problem of competition for these vendors. However, this raises the question; what disadvantages do these vendors have that cause them to lag behind on products? For example, Galleria and T&T are both Asian-fusion grocery chains that rely heavily on foreign products, which could make imports very expensive, potentially explaining the lack of items that have. Without information on the types of products, commenting on the efficiency and competition on these vendors is impossible.

>>>>

---
title: "Grocery stores in Canada"
subtitle: "What vendors hold the most items?"
author: 
  - Hyunje Park, Charlie Zhang
thanks: "Code and data are available at: https://github.com/davidpxrk/cost-of-living-toronto"
date: today
date-format: long
abstract: "This paper analyzes the Canadian Grocery Price data from Canadian grocery vendors, which was gathered by Jacob Filipp and hosted publicly by Project Hammer. It uses information on items, and the name of vendors to determine how many vendors hold a unique grocery item. It was discovered that vendors such as Loblaws, Metro and Walmart held the most unique items at around 20,000 items, while international-focused vendors such as Galleria and T&T held the least amount of items below 10,000."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The `product` table contained 9 variables; most notably `id` and `vendor`. These two variables were important in this analysis, which captured the information on the product id, and the vendor that item was sold at. Data manipulation was conducted through `SQL` [@citesql], to select the variables `id`, `vendor`, and find the number of unique items each vendor sold. This information was exported as a `.csv` file, and the data analysis was conducted through `R Programming Language` [@citeR], `ggplot2` [@citeggplot2],.

@@

The graph above show that vendors such as Metro and Walmart far exceeded the number of products compared to vendors such as Galleria, or T&T, highlighting a problem of competition for these vendors.

## Correlation vs Causation

Larger vendors, tend to have more unique items. Walmart, for example is the biggest chain in the United States, and is also the largest company in the world by revenue, whereas vendors such as Galleria or T&T are much larger in scale. Hence, this could explain the huge difference in item catalogue.

## Missing Data

There were some information on products. For example, there was a variable for `upc`, a universally global code of a product, which was missing for every product, hence this attribute was left out of the analysis. Similarly, `sku`, which is an identifier used by retailers for inventory management was left out due to the same reason.

## Source of bias

However, this raises the question; what disadvantages do these vendors have that cause them to lag behind on products? For example, Galleria and T&T are both Asian-fusion grocery chains that rely heavily on foreign products, which could make imports very expensive, potentially explaining the lack of items that have. Without information on the types of products, commenting on the efficiency and competition on these vendors is impossible.
